# RxODE syntax {#syntax}

This briefly describes the syntax used to define models
that `RxODE` will translate into R-callable compiled code. It also
describes the communication of variables between `R` and the
`RxODE` modeling specification.

## Example

```
   # An RxODE model specification (this line is a comment).

   if(comed==0){   # concomitant medication (con-med)?
      F = 1.0;     # full bioavailability w.o. con-med
   }
   else {
      F = 0.80;    # 20% reduced bioavailability
   }

   C2 = centr/V2;  # concentration in the central compartment
   C3 = peri/V3;   # concentration in the peripheral compartment

   # ODE describing the PK and PD

   d/dt(depot) = -KA*depot;
   d/dt(centr) = F*KA*depot - CL*C2 - Q*C2 + Q*C3;
   d/dt(peri)  =                      Q*C2 - Q*C3;
   d/dt(eff)   = Kin - Kout*(1-C2/(EC50+C2))*eff;
```


## Syntax

```{r echo=FALSE, results='asis'}
cat(unlist(knitr::knit_child(text = gsub("vignette", "section",
                                         gsub("http.*event.*\\.html", "#events", readLines('RxODE/man/rmdhunks/RxODE-syntax-hunk.Rmd'))),
                             envir = environment(), quiet = TRUE)), sep="\n")
```

## cmt() changing compartment numbers for states

The compartment order can be changed with the `cmt()` syntax in the
model.  To understand what the `cmt()` can do you need to understand
how `RxODE` numbers the compartments.

Below is an example of how RxODE numbers compartments

### How RxODE numbers compartments

```{r child='RxODE/man/rmdhunks/rxode-cmt-assign.Rmd'}
```


### Changing compartments by pre-declaring with `cmt()`


```{r child='RxODE/man/rmdhunks/rxode-cmt-preassign-1.Rmd'}
```

### Appending compartments to the model with `cmt()`

```{r child='RxODE/man/rmdhunks/rxode-cmt-preassign-2.Rmd'}
```

